What can GitHub and Googlecode Issues tell us about Exceptions in Android?

### 2014-08-31 android-stacks.tex
all over

This paper reports a repository mining study which analyzed all the issues created on
Android projects available in GitHub (482 applications) and Google code (141 applications)
Overall, our research set includes XXXX issues from which XXX stack traces 
were extracted. Such stack traces were distilled and analyzed in 
combination with source code and bytecode analysis. This study 
consistently identifies common characteristics of reported 
exception stacks traces - e.g., the NullPointerException was the reported cause 
on 30,08\% of all analyzed exception stack traces, and was mentioned in approximately 
50% of all projects analyzed. Moreover,  the cross-project analysis 
of exception stack traces enabled to consistently reveal   
fragilities of exception handling code: 
(1) undocumented runtime exceptions signaled by third party code,
(2) undocumented checked exceptions thrown by Android native code
(3) unexpected exception wrappings (e.g., Errors being wrapped in checked exceptions) 
revealing that Java hybrid exception model is not fully used according to its purpose.
Such fragilities impose challenges to application robustness not only 
in Android environment but also to other systems based on Java exception model.

@@

Such fragilities represent obstacles to the development of dependable java-based applications as they make difficult and 
sometimes almost impossible to protect the code against the Uncaught Exceptions that lead to system crashes.
Although the study focused on Android open-source applications, some of the results are not restricted to this context, 
for instance: some of the unexpected wrappings were performed by Java libraries and elements (e.g. reflection library and 
the static blocks); and the undocumented runtime exceptions were also found in Java libraries besides the Android platform.
The study results are relevant to: (i) Android and Java developers to familiarize common characteristics and fragilities of exception-related code 
which is the first step to help developers preventing and dealing with them; (ii) designers of languages and tools to consider ways of reducing 
such fragilities in Android and Java environments.

@@

 A common way of  propagating exceptions in Java programs is through exception wrapping
 (also known as exception chaining). Exception wrapping allows one exception 
to be wrapped in another one and re-thrown. Figure~\ref{fig:wrapping} presents 
a stack trace associated which illustrates an exception wrapping. The bottom 
part of the stack trace is the \emph{root cause}, which indicates the
first reason for the error to be thrown (in this case, the computer run out of
memory). The top part of the stack trace indicates the location of the exception
manifestation (which will call \emph{top level exception} along this paper). The
execution flow  between the root cause and the exception manifestation may
include other intermediate exception wrappings. In all levels, the exception
\emph{signaler}, is the method that threw the exception, represented on the
stack trace as the first method call below the exception declaration.

@@

The information extracted by each module was structured in a Database,
which contained the distilled information of stack traces, exception types,
and signalers origin (e.g., root exception and its type, root signaler and its origin,
exception wrapper type, wrapper signaler).

@@

>>>>

This paper reports a repository mining study which analyzed all the issues created on
Android projects available in GitHub (482 applications) and Google code (141 applications)
Overall, our research set includes XXXX issues from which XXX stack traces 
were extracted. Such stack traces were distilled and analyzed in 
combination with source code and bytecode analysis. This study 
consistently identifies common characteristics of reported 
exception stacks traces - e.g., the NullPointerException was the reported cause 
on 30,08\% of all analyzed exception stack traces, and was mentioned in approximately 
50\% of all projects analyzed. Moreover,  the cross-project analysis 
of exception stack traces enabled to consistently reveal   
fragilities of exception handling code: 
(1) undocumented runtime exceptions signaled by third party code,
(2) undocumented checked exceptions thrown by Android native code
(3) unexpected exception wrappings (e.g., Errors being wrapped in checked exceptions) 
revealing that Java hybrid exception model is not fully used according to its purpose.
Such fragilities impose challenges to application robustness not only 
in Android environment but also to other systems based on Java exception model.

@@

Such fragilities represent obstacles to the development of dependable java-based applications as they make difficult and 
sometimes almost impossible to protect the code against the Uncaught Exceptions that lead to system crashes.
Although the study focused on Android open-source applications, these findings are not restricted to this context, since 
the above mentioned fragilities may happen in any Java-based system. 
For instance: the unexpected wrappings were performed by Java libraries and elements (e.g. reflection library and 
the static blocks); and the undocumented runtime exceptions were also found in Java libraries besides the Android platform.
The study results are relevant to: (i) Android and Java developers to familiarize common characteristics and fragilities of exception-related code 
which is the first step to help developers preventing and dealing with them; (ii) designers of languages and tools to consider ways of reducing 
such fragilities in Android and Java environments.

@@

 A common way of  propagating exceptions in Java programs is through exception wrapping
 (also known as exception chaining). Exception wrapping allows one exception 
to be wrapped in another one and re-thrown. Figure~\ref{fig:wrapping} presents 
a stack trace  illustrates an exception wrapping. The bottom 
part of the stack trace is the \emph{root cause}, which indicates the
first reason for the error to be thrown (in this case, the computer run out of
memory). The top part of the stack trace indicates the location of the exception
manifestation (which will call \emph{top level wrapper} along this paper). The
execution flow  between the root cause and the exception manifestation may
include other intermediate exception wrappings. In all levels, the exception
\emph{signaler}, is the method that threw the exception, represented on the
stack trace as the first method call below the exception declaration.

@@

The information extracted by each module was structured in a Database,
which contained the distilled information of stack traces, exception types,
and signalers origin (e.g., root cause, root cause type, root cause signaler and its origin,
exception wrappers,  types, and its signalers).

@@

\begin{figure} \centering \includegraphics[scale=0.4]{codeexample2.pdf}
\caption{Code snipets illustrating fragilities on the exception handling code} \label{fig:exchier} \end{figure}
